\label{sec:metrics}

Once defined the coding schemes behavior in terms of encoding
and decoding, we proceed to describe the metrics considered in our study.
We consider the goodput defined as the ratio of the useful delivered
information at the application layer and the total delivery time. We focus
on the goodput considering only the coding process, i.e. we assume that
the application data has been properly generated before encoding and
also correctly post-processed after decoding. The goodput is a measure
for the effective processing speed since it excludes protocol overhead
but considers all delays related with algorithmic procedures, field
operations, hardware processing, multicore coordination (where it applies),
etc. Moreover, encoding and decoding goodput permits to observe if coding is
a system-block that limits the end-to-end performance. As a reference, a low
goodput would affect the \ac{QoE} of delay-intolerant applications for the
end user. For example, Raspberry Pi processors are based in the \ac{ARM}
architecture which is the same as in mobile devices such as smartphones or
tables. Mobile user applications are typically delay-intolerant.
Therefore, the Raspberry Pi might be used as an experiemental tool to
get an estimate of computing capabilities of a mobile device that is
much more easy-deployable and at a lower overall cost.

To complement our study, we review the energy consumption of the Raspberry
Pi since this platform is deployed at a large scale in scenarios where (i)
energy is constrained to the lifetime of the device battery and (ii) the
devices could be established in locations that are unavailable for
regular maintenance. Typical use cases of this type of scenarios are
sensor applications where devices are positioned for measurement retrieval
without any supervision for large periods fo time. In this type of settings,
both overall and per-bit energy consumption of the devices during the
encoding and decoding process are relevant parameters that impact in the
network performance for a given coding scheme.

\subsection{Goodput}
\subsubsection{Encoding}
\subsubsection{Decoding}

\subsection{Energy Consumption}
\subsubsection{Average }
\subsubsection{Per-Bit}